|

### 2025-03-08 output/sandbox_slideshow.qmd
Create final version of space transform choices

## Generalización: Ejes oblicuos

:::::: {.columns .fragment}
::: column
```{r}
x_limits   <- c(-4, 4) # Both limits expressed in canonical metric
y_limits   <- c(-4, 4) # Both limits expressed in canonical metric
break_step <- 2 # "Break units" for both grids, expressed in transformed metric

vec_1 <- c(sqrt(1), sqrt(0))
vec_2 <- c(sqrt(.25), sqrt(.75))
basis <- cbind(vec_1, vec_2) |> unname()

# Basis vectors tangents and norms
vec_tan <- basis[2, ] / basis[1, ]
vec_mod <- basis^2 |> colSums() |> sqrt()

@@

# x_axis_titpos <- -y_limits[1] +
#   if(is.infinite(vec_1_tan)) 0 else x_limits[2] * vec_1_tan
# x_axis_titjust <- x_axis_titpos / x_axis_span
# y_axis_titpos  <- -x_limits[1] +
#   if(is.infinite(vec_2_tan)) 0 else y_limits[2] / vec_2_tan
# y_axis_titjust <- y_axis_titpos / y_axis_span

@@

if (is.infinite(vec_tan[2])) {

  v_grid <- grid_bounds |>
    filter(grid == "v") |>
    mutate(
      unit     = vec_1[1],
      n_units  = trunc(x / unit),
      n_breaks = n_units %/% break_step * break_step
    ) |>
    select(grid, limit, n_breaks, unit) |>
    pivot_wider(names_from = "limit", values_from = n_breaks) |>
    reframe(
      coord   = seq(from = inf, to = sup, by = break_step),
      x_coord = coord * unit
    ) |>
  mutate(label_pos = x_coord, ref = as.character(!coord))

  grid_specs <- grid_specs |> filter(!is.infinite(slope))

  geom_vgrid <- geom_vline(
    mapping = aes(xintercept = x_coord, linetype = ref),
    data = v_grid
  )
}

grid <- grid_specs |>
  select(grid, limit, slope, unit_intercept, n_breaks) |>
  pivot_wider(names_from = limit, values_from = n_breaks) |>
  group_by(grid) |>
  reframe(
    slope,
    coord = seq(from = inf, to = sup, by = break_step),
    intercept = coord * unit_intercept
  ) |>
  mutate(
    label_pos = if_else(
      grid == "h",
       intercept   + x_limits[1] * vec_tan[1],
      (y_limits[1] - intercept)  / vec_tan[2]
    ),
    ref = as.character(!intercept)
  )

scale_linetype <- c(`TRUE` = 'solid', `FALSE` = 'dashed')

@@

grid_plot +
  scale_x_continuous(
    minor_breaks = NULL,
    breaks = grid |> filter(grid == "v") |> pull(label_pos),
    labels = grid |> filter(grid == "v") |> pull(coord),
    limits = x_limits,
    expand = expansion(),
    name   = NULL,
    sec.axis = dup_axis(name = "*&theta;*~2~", labels = NULL)
  ) +
  scale_y_continuous(
    minor_breaks = NULL,
    breaks = grid |> filter(grid == "h") |> pull(label_pos),
    labels = grid |> filter(grid == "h") |> pull(coord),
    limits = y_limits,
    expand = expansion(),
    name   = NULL,
    sec.axis = dup_axis(name = "*&theta;*~1~", labels = NULL)
  ) +
  scale_linetype_manual(values = scale_linetype, guide = NULL) +
  coord_fixed() +
  theme(
    axis.ticks   = element_blank(),
    axis.line    = element_blank(),
    # axis.title   = element_markdown(vjust = y_axis_titjust),
    # axis.title.y = element_markdown(hjust = x_axis_titjust)
    axis.title   = element_markdown(),
    axis.title.y = element_markdown()
  )
```
:::

::: column
:::

::: column
:::
::::::

<!-- # TODO: Opciones de elección de bases oblicuas: rectangular, ejes muy alineados, ejes correctos -->

>>>>

## Generalización: Ejes oblicuos {.smaller}

```{r create-grid-function}
transform_grid <- function(basis,
                           x_limits = c(-4, 4),
                           y_limits = c(-4, 4),
                           break_step = 2,
                           linetype_axis = 'solid',
                           linetype_grid = 'dashed') {
  # Argument checking and formatting: ----

  ## Grid plot linetypes
  linetypes <- c(`TRUE` = linetype_axis, `FALSE` = linetype_grid)

  ## Basis vectors tangents and norms
  vec_tan <- basis[2, ] / basis[1, ]

  ## Axis title positions
  y_axis_titpos <- (y_limits[2] / vec_tan[2] - x_limits[1]) /
    (x_limits[2] - x_limits[1])
  x_axis_titpos <- (x_limits[2] * vec_tan[1] - y_limits[1]) /
    (y_limits[2] - y_limits[1])

  # Main: ----

@@

@@

  # Create vertical grid, if the vertical basis vector is completely vertical
  if (is.infinite(vec_tan[2])) {

    # Create vertical grid data
    v_grid <- grid_bounds |>
      filter(grid == "v") |>
      mutate(
        unit     = basis[1, 1],
        n_units  = trunc(x / unit),
        n_breaks = n_units %/% break_step * break_step
      ) |>
      select(grid, limit, n_breaks, unit) |>
      pivot_wider(names_from = "limit", values_from = n_breaks) |>
      reframe(
        coord   = seq(from = inf, to = sup, by = break_step),
        x_coord = coord * unit
      ) |>
      mutate(label_pos = x_coord, ref = as.character(!coord))

    # Filter out non-valid grid values
    grid_specs <- grid_specs |> filter(!is.infinite(slope))

    # Assign x axis label values and coordinates (for the vertical case)
    x_labels <- v_grid |> select(coord, label_pos)

    # Create vertical grid geometry
    geom_vgrid <- geom_vline(
      mapping = aes(xintercept = x_coord, linetype = ref),
      data = v_grid
    )
  }

  # Create grid data
  grid <- grid_specs |>
    select(grid, limit, slope, unit_intercept, n_breaks) |>
    pivot_wider(names_from = limit, values_from = n_breaks) |>
    group_by(grid) |>
    reframe(
      slope,
      coord = seq(from = inf, to = sup, by = break_step),
      intercept = coord * unit_intercept
    ) |>
    mutate(
      label_pos = if_else(
        grid == "h",
         intercept   + x_limits[1] * vec_tan[1],
        (y_limits[1] - intercept)  / vec_tan[2]
      ),
      ref = as.character(!intercept)
    )

  # Assign x axis label values and coordinates (accounting for the vertical
  #   special case)
  x_labels <- if (is.infinite(vec_tan[2]))    v_grid |> select(coord, label_pos)
              else       grid |> filter(grid == "v") |> select(coord, label_pos)

@@

  grid_plot +
    scale_x_continuous(
      minor_breaks = NULL,
      breaks = x_labels |> pull(label_pos),
      labels = x_labels |> pull(coord),
      limits = x_limits,
      expand = expansion(),
      name   = NULL,
      sec.axis = dup_axis(name = "*&theta;*~2~", labels = NULL)
    ) +
    scale_y_continuous(
      minor_breaks = NULL,
      breaks = grid |> filter(grid == "h") |> pull(label_pos),
      labels = grid |> filter(grid == "h") |> pull(coord),
      limits = y_limits,
      expand = expansion(),
      name   = NULL,
      sec.axis = dup_axis(name = "*&theta;*~1~", labels = NULL)
    ) +
    scale_linetype_manual(values = linetypes, guide = NULL) +
    coord_fixed() +
    theme(
      axis.ticks   = element_blank(),
      axis.line    = element_blank(),
      axis.title.x       = element_markdown(hjust = y_axis_titpos),
      axis.title.y.right = element_markdown(vjust = x_axis_titpos)
    )
}
```

<br>

:::::: columns
::: {.column .fragment width="33%"}
<center>A</center>

```{r grid-orthogonal-rotation}
#| fig-width: 5
basis <- cbind(
  c(sqrt(.85), -sqrt(.15)),
  c(sqrt(.15), sqrt(.85))
)

transform_grid(basis)
```
:::

::: {.column .fragment width="33%"}
<center>B</center>

```{r grid-ts-basis}
#| fig-width: 5
basis <- cbind(
  c(1, 0),
  c(sqrt(.25), sqrt(.75))
)

transform_grid(basis)
```
:::

::: {.column .fragment width="33%"}
<center>C</center>

```{r grid-ls-basis}
#| fig-width: 5
basis <- t(solve(basis))

transform_grid(basis)
```
:::
::::::
